---
title: "Urban Accidents in the City of Porto Alegre"
author: "Jean-Marc Vincent, Lucas Mello Schnorr"
date: "October 2017"
output:
  pdf_document:
    number_sections: yes
  html_document: default
geometry: margin=1.5in, top=0.5in, bottom=0.5in
---

Each student should provide a Rmd file with _two_ to _four_ plots, with text describing the semantics of the data, the question, how they have answered the question, and an explanation for each figure, showing how that particular figure helps the answering of the initial question.  Fork the LPS repository in GitHub, push your Rmd solution there. Send us, by e-mail, the link for your GIT repository, indicating the PATH to the Rmd file. Check the LPS website for the deadline.

# Introduction

The City of Porto Alegre, under the transparency law, has provided a data set with all the urban accidents (within the city limits) since 2000. The data set, including a description of each column in the PDF file format, is available in the following website:

http://www.datapoa.com.br/dataset/acidentes-de-transito

# Goal

For a given year (defined by the LPS coordination for each student enrolled in the cursus), the goal is to answer one of the following questions. The solution must use the data import and manipulation verbs of the R programming language and the tidyverse metapackage (readr, tidyr, dplyr) using Literate Programming.

# Questions

1. What is the time of the day with most accidents?
2. How many vehicles are involved in the accidents?
3. What types of accidents are more common?
4. Is the number of deaths increasing or decreasing?
5. Is there a street of the city with more accidents than others?
6. Do holidays impact in the number of accidents?

# Download the data

Supposing you have the URL for the CSV file, you can read the data using the code below. You can also download it manually and commit it to your repository to avoid an internet connection every time you knit this file. If the URL changes, the second solution might even make your analysis be more portable in time.

```{r}
library(readr)
#URL <- "http://www.opendatapoa.com.br/storage/f/2013-11-06T16%3A52%3A35.356Z/acidentes-2009.csv"
#URL <- "http://www.opendatapoa.com.br/storage/f/2013-11-06T16%3A38%3A56.849Z/acidentes-2007.csv"

#df <- read_delim(URL, delim=";")
df <- read_delim("acidentes-2007.csv", delim=";")
df
```
# 1 - What is the time of the day with most accidents?

```{r}
library(magrittr);
library(dplyr);

df %>% group_by (NOITE_DIA) %>% summarize(TOTAL = n());
```

The amount of accidents during the day is over twice the nightly amount, which is kind of unexpected. Although there tend to be more cars in the streets during the day, at night it's probably easier for one to become distracted/sleepy/drunk driving.


```{r}
df %>% group_by(TIPO_ACID, NOITE_DIA) %>% summarize(TOTAL = n())
```

Checking the amount of accidents of each kind we see that the only one that happens more at night than during the day is CAPOTAGEM (overturning), which doesn't necessarily mean anything but could suggest that it's easier to flip your car at night.

# 2 - How many vehicles are involved in the accidents?

```{r}

df %>% summarize(TOTAL_AUTO = sum(AUTO),
               TOTAL_TAXI = sum(TAXI),
               TOTAL_LOTACAO = sum(LOTACAO),
               TOTAL_ONIBUS_URB = sum(ONIBUS_URB),
               TOTAL_ONIBUS_INT = sum(ONIBUS_INT),
               TOTAL_CAMINHAO = sum(CAMINHAO),
               TOTAL_MOTO = sum(MOTO),
               TOTAL_CARROCA = sum(CARROCA),
               TOTAL_BICICLETA = sum(BICICLETA),
               TOTAL_OUTRO = sum(OUTRO)) %>% t() %>% as.data.frame %>% rename(Amount = V1) # %>% arrange(desc(Amount))

```

The table was transposed to become easier to read its results. As we can see, the majority of accidents is caused by autos (which probably means cars, but I can't be sure), followed by motos (bikes). This result suggests that the worst drivers we have around here are the "common people". It's also interesting to note that ONIBUS_URB (urban buses) are the most involved in accidents, so it could be safer going from another city to Porto Alegre (via ONIBUS_INT - intermunicipal buses [and I'm just guessing actually]) than traveling inside the city, unless you take a LOTACAO (I don't know if there is an english name for this kind of bus).

PS: It would be better if the frame was ordered by the amount of accidents, but everytime I tried arranging it, the row descriptions were gone so I just left it as it is.

# 3. What types of accidents are more common?

```{r}
df %>% group_by(TIPO_ACID) %>% summarize(Amount = n()) %>% arrange(desc(Amount))

```

The most common is by far ABALROAMENTO, which, according to Google, is "a violent shock of one thing against another". This description turns difficult to analyze this data, cause I fail to see a difference in comparison to "CHOQUE" (shock).

```{r}
df %>% group_by(TIPO_ACID) %>% summarize(Deaths = sum(MORTES),
                                          Hurtings = sum(FERIDOS)) %>% arrange(desc(Deaths, Hurtings))
```

Checking for the amount of hurt/dead people involved in those accidents, we see that, although CHOQUE and ABALROAMENTO seems very similar, the former tends to kill people while the latter tends to only hurt them. So, maybe CHOQUE stands for more violent incidents (like two cars colliding) and ABALROAMENTO for any other type of collision.

# 4. Is the number of deaths increasing or decreasing?

```{r}
df %>% group_by(MES) %>% summarize(Mortes = sum(MORTES)) %>% plot(type = "l")
```

In the graph above, we can see that the most life-threatening months are July and December. This might be because of the holydays, when a lot of people travel and possibly expose themselves to this kind of danger.

```{r}
df %>% group_by(MES) %>% summarize(Mortes = sum(MORTE_POST)) %>% plot(type = "l")
```

Checking for posthumous death, the graph isn't that different, except that it now has a lot of incidents on October also.

# 5. Is there a street of the city with more accidents than others?

```{r}
df %>% group_by(LOG1) %>% summarize(Accidents = n()) %>% arrange(desc(Accidents))
```

The majority of accidents happens in avenues, which is completely expected, since they are the most frequented streets.

```{r}
df %>% filter(LOG1 == "AV IPIRANGA") %>% group_by(NOITE_DIA) %>% summarize(n())
```

Checking AV IPIRANGA, the most dangerous avenue (and part of the way to Campus do Vale to many people), we see that the accidents tend to happen during the day (when we are going to the campus!), so it's not a night-time problem or something like that.






